---
title: "Activity Monitoring Data"
author: "Namakuto"
date: "June 25, 2017"
output: 
  html_document: 
    keep_md: yes
---

In this R Markdown document, activity monitoring data on daily walking step counts as well as the associated time intervals for said activity were recorded. 
This document will echo the code used to answer the questions of Week 2 in the Reproducible Research course (as taught by John Hopkins).

___
### Loading and preprocessing the data
```{r loading, echo=TRUE, results='asis',warning=FALSE}
mydata<-read.csv("activity.csv")
attach(mydata)
library(ggplot2)
mydata.split<-split(mydata,f=date)
```

___
### What is mean total number of steps taken per day?
```{r plot, echo=FALSE, warning=FALSE, results='Asis'}
sum.split<-sapply(mydata.split, function(x) sum(x[[1]]))
qplot(sum.split, geom="histogram",
      main="Histogram of Daily Step Counts, Summed", 
      xlab="Steps/Day, Summed")
```

Evaluation of the mean and median number of steps/day:
```{r finding the mean and median, echo=FALSE}
newdata.sums<-data.frame(levels(date),unname(sum.split))
colnames(newdata.sums)<-c("Date", "Total # of Steps")
cat("Mean is: ", mean(newdata.sums$`Total # of Steps`, na.rm=TRUE))
cat("Median is: ", median(newdata.sums$`Total # of Steps`, na.rm=TRUE))
```

___
### What is the average daily activity pattern?
```{r line-graph, echo=FALSE}
mydata.split_i<-split(mydata, f=factor(interval)) 
mean.split_i<-sapply(mydata.split_i, function(x) mean(x[,1], na.rm = TRUE))

f.interval <-levels(factor(interval))
plot(f.interval,mean.split_i, type="l", lwd=1, col="purple",
     main="Mean Number of Steps per 5-Minute Interval",
     ylab="Steps", xlab="Interval (Min)", las=1, xaxt="n")
axis(1,at=seq(0,2400,100),cex.axis=0.8)
```

Which 5-min interval contains the max number of steps?
```{r max,echo=FALSE}
row<-grep(pattern=max(mean.split_i),mean.split_i,value=FALSE)
cat("Interval point: ", f.interval[104])
```

___
### Imputing Missing values
Number of "Na's" in the dataset:
```{r missing values, echo=FALSE}
bad<-mydata[is.na(mydata)]
cat("NA count: ", length(bad))
```

Replacement of "NA's", imputted back into the original dataset:
```{r replacing missing values, echo=TRUE, results='asis'}
mydata[is.na(mydata)]<-0
```

```{r non-NA-bar-chart, echo=FALSE, results='Asis', warning=FALSE}
sum.split<-sapply(mydata.split, function(x) sum(x[[1]]))
qplot(sum.split, geom="histogram",
      main="Histogram of Daily Step Counts, Na's changed to 0", 
      xlab="Steps/Day, Summed")
```

There does not appear to be a significant difference in replacing the "NA's" with "0's".

___
### Are there differences in activity patterns between weekdays and weekends?

```{r activity-in-week, echo=FALSE, warning=FALSE}
mydata[is.na(mydata)]<-0
mydata.weekdays<-weekdays(as.Date(mydata$date))
mydata$date<-mydata.weekdays

mydata$date<-gsub(pattern="Monday", replacement="Weekday", mydata$date)
mydata$date<-gsub(pattern="Tuesday", replacement="Weekday", mydata$date)
mydata$date<-gsub(pattern="Wednesday", replacement="Weekday", mydata$date)
mydata$date<-gsub(pattern="Thursday", replacement="Weekday", mydata$date)
mydata$date<-gsub(pattern="Friday", replacement="Weekday", mydata$date)
mydata$date<-gsub(pattern="Saturday", replacement="Weekend", mydata$date)
mydata$date<-gsub(pattern="Sunday", replacement="Weekend", mydata$date)

daytype.split<-split(mydata,f=factor(mydata$date)) # new data.frame with $week
daytype.split1<-daytype.split$Weekday
daytype.split2<-daytype.split$Weekend

mydata.split_i1<-split(daytype.split1, f=factor(interval))
mydata.split_i2<-split(daytype.split2, f=factor(interval))

means.daytype1<-sapply(mydata.split_i1, function(x) mean(x[,1]))
means.daytype2<-sapply(mydata.split_i2, function(x) mean(x[,1]))
par(mfrow=c(2,1),mar=c(2.5, 4 ,1.5, 1 ))
f.interval <-levels(factor(interval))
plot(f.interval,means.daytype1, type="l", lwd=1, col="dark cyan",
     main="Mean Number of Steps per 5-Minute Interval, Weekdays",
     ylab="Steps", xlab="Interval (Min)", las=1, xaxt="n", cex.main=1)
axis(1,at=seq(0,2400,100),cex.axis=0.8)
plot(f.interval,means.daytype2, type="l", lwd=1, col="dark cyan",
     main="Mean Number of Steps per 5-Minute Interval, Weekends",
     ylab="Steps", xlab="Interval (Min)", las=1, xaxt="n", cex.main=1)
axis(1,at=seq(0,2400,100),cex.axis=0.8)
```

There appears to be some difference between weekday and weekend activity patterns, with greater peaks in the latter's data. Weekday activity has a higher maximum step count, however. 
